---
title: "Wrangle rows and columns"
format: html
---

# Instructions

This week, you learned several more functions from the dplyr package (pronounced DEE-ply-er). These functions help you work with the rows and columns in your data set.

You'll use these skills to determine if Global Health (GH) priority countries see a larger increase in children's health than other countries.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rstudio.github.io/academy-tidyverse-foreign-aid/assets/j2sr_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Create and select columns

In this milestone, you'll create a new variable that indicates whether or not a country received priority aid as a USAID Global Health (GH) priority country. You'll then determine if priority countries see a larger increase in children's health than other countries.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `j2sr.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `j2sr`.

```{r}
#| label: recreation-import

```

### Part 2 - Prepare the data

Run the code below to recreate the table of children's health change we built last week. You will use this table to create and explore a new variable.

```{r}
#| label: prep-data
health_change <-
  j2sr |>
  filter(year %in% c(2014, 2020)) |>
  group_by(country) |>
  summarize(child_health_change = last(child_health) - first(child_health))

health_change
```

### Part 3 - Create a new column

We would like to add a new column that records whether or not each country was a USAID GH priority country. Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone03.csv")
solution
```

Your task is to use what you've learned about the dplyr package to transform `health_change` into this table. You will need to:

1. Create a new column named `priority_country`, which indicates whether or not a country was a USAID GH priority country (given in the `priority_countries` vector below).

```{r}
#| label: create-vector
priority_countries <- c("Ethiopia", "Mozambique", "Nigeria", "Tanzania", "Uganda", "DR Congo", "Ghana", "Malawi", "Mali", "Senegal", "Kenya", "Rwanda", "Zambia", "Liberia", "Madagascar", "Bangladesh", "Nepal", "Burma", "India", "Haiti", "Afghanistan", "Pakistan", "South Sudan", "Yemen", "Burkina Faso", "Niger", "Indonesia", "Cambodia", "Cameron", "Cote d'Ivoire", "Sierra Leone", "Zimbabwe", "Angola", "Guinea", "Thailand", "Benin", "Philippines")
```

Work in the code chunk below. Save the result as `health_priority`.

```{r}
#| label: recreation-create

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(health_priority, solution, tolerance = 1e-4)
```

### Part 4 - Visualize

Now that you have created the `priority_country` column, use it to visualize the distribution of change in child health between 2014 and 2020 among GH priority countries, and other countries.

```{r}
#| label: recreation-visualize

```

Given the results shown in your plot, do you think that GH priority countries experienced a greater increase in children's health than other countries?

## Extension

Using the code chunk below, investigate a single research question about this data, using the additional data wrangling skills you learned this week. Some ideas:

1. How do USAID Global Health priority countries compare to all other countries on average in terms of characteristics such as poverty rate, education quality, GDP per capita, etc.? Looking at these (and other) metrics by country, are you able to identify other countries that could be considered for priority status?
2. Classify countries based on the change in child health scores between 2014 and 2020, using thresholds and categories of your choosing (e.g. "improved", "stagnant", and "declined"). What are the characteristics of countries that fall into each of these categories?
3. Based on the degree of economic gender gaps and/or political equality across social groups, divide this data into groups of your choosing that represent the level of social quality (e.g. low/medium/high, above/below a regional average, etc.) and explore patterns based on these groups.
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Create at least one new variable in your data set (using `mutate()` and/or `summarize()`), or use other dplyr functions to alter your data
3. Use your updated data set to create at least one graph and/or table

```{r}
#| label: extension

```
