---
title: "Work with data types"
format: html
---

# Instructions

This week, you worked with three new data types in R -- factors, strings, and date-times -- using the forcats, stringr, and lubridate packages, respectively.

In this milestone, you'll use factor reordering to visualize how changes in child health over time vary by country and region.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rstudio.github.io/academy-tidyverse-foreign-aid/assets/j2sr_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Data types

In this milestone, you'll use the forcats package along with ggplot2 to visualize how changes in child health over time vary geographically.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `j2sr.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `j2sr`.

```{r}
#| label: recreation-import

```

### Part 2 - Prepare the data

Run the code below to see a table. Your task is to combine code from previous milestones to calculate the change in child health scores from 2014 to 2020 for each country and region.

```{r}
#| label: recreate-this-1
#| message: false
solution <- readr::read_csv("data/milestone05.csv")
solution
```

To create this table, you will need to:

1. Filter the data to only include observations from 2014 and 2020
2. Group the data by `country` and `region`
3. Summarize the `child_health_change` for each country
4. Add the `priority_country` column (using the `priority_countries` vector below)
5. Remove any rows with a missing value of `child_health_change`

```{r}
#| label: create-vector
priority_countries <- c("Ethiopia", "Mozambique", "Nigeria", "Tanzania", "Uganda", "DR Congo", "Ghana", "Malawi", "Mali", "Senegal", "Kenya", "Rwanda", "Zambia", "Liberia", "Madagascar", "Bangladesh", "Nepal", "Burma", "India", "Haiti", "Afghanistan", "Pakistan", "South Sudan", "Yemen", "Burkina Faso", "Niger", "Indonesia", "Cambodia", "Cameron", "Cote d'Ivoire", "Sierra Leone", "Zimbabwe", "Angola", "Guinea", "Thailand", "Benin", "Philippines")
```

Work in the code chunk below. Save the result as `health_change`.

```{r}
#| label: recreation-prep

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(health_change, solution, tolerance = 1e-4)
```

### Part 3 - The role of factor orders in visualizations

You can use the `health_change` table to visualize how the change in children's health varies by country and region. Run the chunk below to view a first attempt at this plot.

```{r}
#| label: first-plot
#| eval: false
health_change |>
  ggplot(aes(x = child_health_change, y = country, fill = priority_country)) +
  geom_col() +
  facet_wrap(vars(region), scales = "free_y") +
  labs(
    x = "Change in Child Health",
    y = "Country"
  )
```

Because the order of the countries on the y-axis and the regions (facets) is arbitrary (alphabetical), it is difficult to glean information from this plot. Let's improve this plot by giving these variables a more principled order.

#### Visualization

Using the File Explorer, navigate to the `images/` directory, then click on `milestone05.png` to open an image.

You should see a new version of this plot in which *both* the order of the countries on the y-axis and the regions (facets) now corresponds to the average change in children's health scores. Your task is to recreate this plot.

Use the chunk below to recreate this plot. Hint: which function from the forcats package can you use to reorder a factor by another variable?

```{r}
#| label: recreation-visualize

```

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. How do total disbursements vary by country and region? How do priority countries compare to non-priority countries in terms of disbursements? Use faceting and factor reordering as you did in your recreation to visualize this.
2. Are disbursements to health-related sectors associated with larger changes in child health compared to disbursements to other types of sectors? Use string matching to wrangle the `d2r` data set in order to explore this question.
3. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. If your data contains ...
  ... date-time data, use a lubridate function to correctly format your date-time variables
  ... string data, use a stringr function to manipulate your string variables
  ... categorical data, use a forcats function to modify the factor levels of your categorical variables (reorder, recode, etc.)
3. Using a table or graph, present a meaningful subset of your modified data

```{r}
#| label: extension

```
