---
title: "Tidy and join tables"
format: html
---

# Instructions

This week you learned about tidying and joining data sets with the tidyr (pronounced tidy-R) and dplyr packages.

Joining data sets can be deceptively tricky. It is important to keep in mind that prepping and inspecting the tables you're working with *before* you do a join is often more important (and takes more time) than doing the join itself.

You'll use some of these skills in the milestone to the right.

To learn more about the variables contained in the data sets, refer to the data dictionaries: https://rstudio.github.io/academy-tidyverse-foreign-aid/assets/j2sr_dictionary.html and https://rstudio.github.io/academy-tidyverse-foreign-aid/assets/d2r_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Tidy and join

In this milestone, you'll check whether larger aid disbursements are associated with larger health improvements. To do this, you'll need to use the Dollars to Results data (in `d2r.csv`) alongside the Journey to Self-Reliance data you've been using (in `j2sr.csv`). Both data sets are stored in your `data/` directory.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data sets. Use the code chunk below to read two data sets from the `data/` directory:

1. `j2sr.csv`, save it as `j2sr`
2. `d2r.csv`, save it as `d2r`

```{r}
#| label: recreation-import

```

### Part 2 - Prepare the data

Run the code below to recreate the table of children's health change we built in previous milestones. You will use this table to join to a new data set.

```{r}
#| label: prep-data
health_change <-
  j2sr |>
  filter(year %in% c(2014, 2020)) |>
  group_by(country) |>
  summarize(child_health_change = last(child_health) - first(child_health))

health_change
```

Next, use what you've learned about the dplyr package to transform `d2r` into a table of total disbursements per country. You will need to:

1. Filter the data to only include disbursements (`result_name == "disbursements"`) between the fiscal years 2014 and 2020
2. Group the data by `country_name`
3. Summarize the `total_disbursements` for each country

Work in the code chunk below. Save the result as `total_disbursements`.

```{r}
#| label: recreation-summary

```

### Part 3 - Join tables

Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone04.csv")
solution
```

Your task is to use what you've learned about the dplyr package to transform `health_change` and `total_disbursements` into this table. You will need to do some pre-join checks:

1. Which variables can you join on?
2. Do the data sets have any mismatched observations?
3. Perform any necessary data cleaning.
4. Check again as needed!

Then, join the data sets so that every country in `health_change` keeps its `child_health_change`, with the matching `total_disbursements` added alongside.

Work in the code chunk below. Save the result as `disbursement_health_change`.

```{r}
#| label: recreation-join

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(disbursement_health_change, solution, tolerance = 1e-4)
```

## Extension

Using the code chunk below, investigate a single research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Create a visualization that takes advantage of the new variable(s) you now have access to. You might find the `slice()` family of functions useful to identify and filter for the top N countries by aid disbursements: https://dplyr.tidyverse.org/reference/slice.html
2. The `d2r` data set contains information on many other "results" besides aid disbursements; explore another metric of interest in this data set and ask a question about it using tables and graphs. You may consider first narrowing your exploration based on sector name (e.g. "Basic Education" or "Water Supply and Sanitation".)
3. Many other country-level data sets may be relevant to this project. Find another country-level variable from a publicly available data set, join in that data, and ask a question of the new data set. (e.g. Is there a relationship between child health and population density?)
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Complete a data wrangling task relevant to your data (tidying and/or reshaping/pivoting your data); if appropriate, join your data with a second data set (filtering or mutating join)
3. Show summary tables of your data throughout these transformations


```{r}
#| label: extension

```
