---
title: "Summarize data"
format: html
---

# Instructions

This week, you learned several new functions from dplyr (pronounced DEE-ply-er), a commonly-used package from the tidyverse for wrangling data. You also learned some new ways to enhance your plots.

You'll use these skills to explore how children's health scores differ across regions, and to compute the change in children's health between 2014 and 2020 for each country.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/foreign-aid/assets/j2sr_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Summarize data

In this milestone you'll create and compare summary statistics for different regions and countries.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `j2sr.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `j2sr`.

```{r}
#| label: recreation-import

```

### Part 2 - Visualize summaries by region

First, let's visually explore whether the distribution of `child_health` differs across regions in 2020. Run the chunk below to view a summary table of the average `child_health` score for each region.

```{r}
#| label: summary
#| eval: false
j2sr |>
  filter(year == 2020) |>
  group_by(region) |>
  summarize(avg_child_health = mean(child_health, na.rm = TRUE))
```

Run the chunk below to visualize the distribution of `child_health` scores across regions in 2020.

```{r}
#| label: summary-plot
#| eval: false
j2sr |>
  filter(year == 2020) |>
  ggplot(aes(x = child_health)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(vars(region))
```

### Part 3 - Summarize change by country

Now, let's compute the change in children's health between 2014 and 2020 for each country. Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone02.csv")
solution
```

Your task is to use what you've learned about the dplyr package to transform `j2sr` into this table. You will need to:

1. Filter the data to only include observations from 2014 and 2020
2. Group the data by country
3. Summarize the `child_health_change` for each country

*Hint:* the `first()` and `last()` functions may prove useful for this type of summary.

Work in the code chunk below. Save the result as `health_change`.

```{r}
#| label: recreation-summary

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(health_change, solution, tolerance = 1e-4)
```

## Extension

Using the code chunk below, investigate a single research question about this data, using the data wrangling skills you learned this week. Some ideas:

1. Data visualization is often an easier way to see patterns than numbers in a table. Build a graph from your summary table of the change in child health between 2014 and 2020 for each country. (Consider using filtering, aesthetics and/or faceting to make this plot easier to interpret.)
2. How does child health vary by income group?
3. Which year had the highest average child health score overall? The lowest? Is this pattern consistent across regions?
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Use `group_by()` and `summarize()` to produce simple a table from some subset of your data
3. Create a related graph of this data subset using ggplot


```{r}
#| label: extension

```
